Off the beaten path: \\ ...

### 2018-04-28 paper/gd1.tex
intro

Over 30 candidate stellar streams have been discovered throughout the halo of
the Milky Way, largely thanks to large-area, multi-band photometric surveys
(especially the Sloan Digital Sky Survey, SDSS; \citealt{York:2000}).
The Milky Way streams have a wide range of properties such as length, velocity
dispersion, and density (e.g., \citealt{Odenkirchen:2001, Grillmair:2006,
Grillmair:2006b, Belokurov:2006, Belokurov:2007, Bonaca:2012, Shipp:2018}; see
also the recent review \citealt{Grillmair:2016, Newberg:2016}), and the majority
have no obvious, bound progenitor systems;
of the thin, likely globular-cluster-origin streams, only two have clear
progenitors (NGC 5466 and Palomar 5).

The most prominent and longest (in angle) thin stream is the GD-1 stream, named
after its discoverers (\citealt{Grillmair:2006}).
GD-1 was found using a matched filter on photometric data from the SDSS and
spans over $60^\circ$ in Heliocentric sky coordinates, partially because of its
relative closeness to the sun ($\approx 7$--$10~\textrm{kpc}$).
Despite its proximity, no remnant progenitor cluster or stellar system has been
found associated with the stream, but its metallicity ($\feh \approx -1.4$;
\citealt{Koposov:2010}) and estimated stellar mass ($M \approx 2 \times
10^4~\msun$; \citealt{Koposov:2010}) are consistent with being a disrupted
globular cluster.

The long physical length of the GD-1 stream ($\sim 15~\kpc$) and location in the
Galactic halo (pericentric distance $r_\textrm{peri} \sim 13~\kpc$) make it an
ideal object for constraining the gravitational potential of dark matter in the
inner Milky Way halo.
Indeed, both from fitting orbits to binned phase-space measurements along the
GD-1 stream (\citealt{Koposov:2010}) and from modeling the stream track in
phase-space coordinates (\citealt{Bovy:2016}), it has been shown that the dark
matter distribution within its orbit is consistent with or nearly spherical.

Its long length --- and therefore its implied old dynamical age --- combined
with its particular orbit also make it a prime stream to search for interactions
with dark matter subhalos:
because of its large pericentric distance and retrograde orbit with respect to
the Galactic bar, density variations in the stream are not expected from either
interactions with giant molecular clouds (\citealt{Amorisco:2016}) or resonant
encounters with the bar (\citealt{Pearson:2017}).
[Erkal 2016 predict ~1 gap]

From SDSS photometry, hints of gaps and under-densities have been identified in
the GD-1 stream, but with low significance (\citealt{Carlberg:2013}).
More recently, deep photometry of GD-1 from CFHT/Megacam has revealed
interesting small-scale ``wiggles'' and significant density variations along a $\approx 45^\citc$ span of the stream (\citealt{DeBoer:2017}):
[summarize their conclusions].

>>>>

Over 30 candidate stellar streams have been discovered throughout the halo of
the Milky Way, thanks to large-area, multi-band photometric surveys (especially
the Sloan Digital Sky Survey, SDSS; \citealt{York:2000}).
The Milky Way streams have a wide range of properties such as length, velocity
dispersion, and density (e.g., \citealt{Odenkirchen:2001, Grillmair:2006,
Grillmair:2006b, Belokurov:2006, Belokurov:2007, Bonaca:2012, Shipp:2018}; see
also the recent review \citealt{Grillmair:2016, Newberg:2016}), and the majority
have no obvious, bound progenitor systems;
of the thin, likely globular-cluster-origin streams, only two have clear
progenitors (NGC 5466 and Palomar 5).

The most prominent and longest (in angular extent) thin stream is the GD-1
stream, named after its discoverers (\citealt{Grillmair:2006}).
GD-1 was found using a matched filter on photometric data from the SDSS and
spans over $60^\circ$ in Heliocentric sky coordinates, partially because of its
relative closeness to the sun ($\approx 7$--$10~\textrm{kpc}$).
Despite its proximity, no remnant progenitor cluster or stellar system has been
found associated with the stream, but its metallicity ($\feh \approx -1.4$;
\citealt{Koposov:2010}) and estimated stellar mass ($M \approx 2 \times
10^4~\msun$; \citealt{Koposov:2010}) are consistent with being a disrupted
globular cluster.

The long physical length of the GD-1 stream ($\sim 15~\kpc$) and location in the
Galactic halo (pericentric distance $r_\textrm{peri} \sim 13~\kpc$) make it an
ideal object for constraining the gravitational potential of dark matter in the
inner Milky Way.
Indeed, both from fitting orbits to binned phase-space measurements along the
GD-1 stream (\citealt{Koposov:2010}) and from modeling the stream track in
phase-space coordinates (\citealt{Bovy:2016}), it has been shown that the dark
matter distribution within its orbit is consistent with or nearly spherical.

Its long length --- and therefore its implied old dynamical age --- combined
with its particular orbit also make it a prime stream to search for interactions
with dark matter subhalos:
because of its large pericentric distance and retrograde orbit with respect to
the Galactic bar, density variations in the stream are not expected from either
interactions with giant molecular clouds (\citealt{Amorisco:2016}) or resonant
encounters with the bar (\citealt{Pearson:2017}).
By assuming an age [...], \citet{Erkal:2016} predict that XX (~1 gap??).

From SDSS photometry, hints of further gaps and under-densities have been
identified in the GD-1 stream, but with low significance
(\citealt{Carlberg:2013}).
More recently, deep photometry of GD-1 from CFHT/Megacam has revealed
interesting small-scale ``wiggles'' and significant density variations along a $\approx 45^\citc$ span of the stream (\citealt{DeBoer:2017}):
[summarize their conclusions].
